Adapting bootstrap to our problem - Part 2

### 2024-07-11 adapt_boot2.Rmd
Better comparison

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
library(ggplot2)
```

@@

  df_boot <- tibble(.rows = nrow(df_sample))
  for (b in 1:nboot) {
    nb <- sample(1:nrow(df_sample), nrow(df_sample), replace = TRUE, prob = df_sample$prob)
    df_sample$prob <- NULL
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with nboot samples of 2 variables and of size n
}
```

@@

```{r}
comparison2 <- function(B) {
  # B: number of bootstrap replications each time
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100
  while (b1  < B) {
    ## First method
    # We only keep x and y in the dataframe
    random_points_1 <- random_points
    random_points_1$l <- NULL
    df_boot <- f_boot(random_points_1, b1)
    est_boot_1 <- est_boot1(df_boot)

    ## Second method
    df_boot2 <- f_boot2(random_points, b1)
    est_boot_2 <- est_boot1(df_boot2)

@@

    # tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    # tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b1 <- b1+10
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

```{r}
res2 <- comparison2(500)
```

```{r}
res <- comparison2(500)
```

```{r}
data = as.data.frame(cbind(1:length(res2[[1]]), res2[[1]], res2[[2]]))
p1 <- ggplot(data, aes(data[,1], data[,2]))
p2 <- ggplot(data, aes(data[,1], data[,3]))
p1 + p2 + geom_point()
p1 + geom_point

# p <- ggplot(data, aes(data[,1], cbind(data[,2], data[,3])))
# p + geom_point()

p <- ggplot(data, aes(data[,1], data[,3] - data[,2]))
p + geom_point()
```

@@

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

```{r}
ggplot(x = 1:length(res[[1]]), y = res[[1]]) +
  plot(x = 1:length(res[[1]]), y = res[[2]], col = "red")
geom_point(
    mapping = aes(y = group_mean), data = group_means_df,
    colour = 'red', size = 3)

```

```{r}
# If it is positive, the second method is the best to minimize the variance
sum(res[[2]])
# With ncomp = 100 and B = 500
# 2 e.g.
# 12

# With ncomp = 1000 and B = 800
# 56
```

# Conclusion

>>>>

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
library(ggplot2) # To plot scatterplots
library(dplyr) # for select
```

@@

  df_boot <- tibble(.rows = nrow(df_sample))
  for (b in 1:nboot) {
    nb <- sample(1:nrow(df_sample), nrow(df_sample), replace = TRUE, prob = df_sample$prob)
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  df_boot <- do.call(data.frame, df_boot)
  df_boot <- df_boot %>% select(-contains("prob")) # Probabilities are useless now
  return(df_boot)
  # Return a data frame with nboot samples of 2 variables (x,y,local variable) and of size n
}
```

@@

```{r}
comparison2 <- function(df_init, B, small_step) {
  # df_init: initial data frame with 4 variables, just as random_points
  # B: number of bootstrap replications each time
  # small_step: step to decide how many bootstrap series will be computed
  # large small_step => a few bootstrap series
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100
  while (b1  < B) {
    ## First method
    # We only keep x and y in the dataframe
    df_init_1 <- df_init
    df_init_1$l <- NULL
    df_boot <- f_boot(df_init_1, b1)
    est_boot_1 <- est_boot1(df_boot)

    ## Second method
    df_boot2 <- f_boot2(df_init, b1)
    est_boot_2 <- est_boot1(df_boot2)

@@

    # tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    # tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b1 <- b1+small_step
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

```{r}
B_test <- 800
```

```{r}
df_test <- random_points
res2 <- comparison2(df_test, B_test, 1)
```

```{r}
# We look at the mean of estimators
data = as.data.frame(cbind(1:length(res2[[1]]), res2[[1]], res2[[2]]))

p <- ggplot(data, aes(data[,1], data[,2] - data[,3]))
p + geom_point() +ggtitle("Comparison of the mean of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")
```

@@

p <- ggplot(data_var, aes(data_var[,1], data_var[,2] - data_var[,3]))
p + geom_point() + ggtitle("Comparison of the variance of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")

# If it is positive, method 2 is better
```

```{r}
# To what extent does the second method improve variance?
cat("If it is positive, it indicates that the second method improve reducing the variance.")
cat("\nWith all bootstrap samples:", mean(data_var[,2] - data_var[,3]))
cat("\nWith the first half of bootstrap series (between 100 and ", floor((B_test+100)/2)," samples): ", mean(data_var[1:(floor((B_test+100)/2)-100),2] - data_var[1:(floor((B_test+100)/2)-100),3]), sep = "")
cat("\nWith the last half of bootstrap series (between ", floor((B_test+100)/2)," and ", B_test, " samples): ", mean(data_var[(floor((B_test+100)/2)-100):(B_test-100),2] - data_var[(floor((B_test+100)/2)-100):(B_test-100),3]), sep = "")

# Ex:
# With all bootstrap samples: 13172.49
# With the first half of bootstrap series (between 100 and 300): -3856.388
# With the last half of bootstrap series (between 300 and 500 samples): 27526.69
```
